
\subsection{Normal model built on Probit representation of Spherical Coordinate Space}
\label{method:npprobitnorm}
The transformation in Equation~\ref{eqn:transform} provides us a mapping from $S_{2}^{d-1}$ to a
  $d-1$ dimensional cube, $[0, \pi/2]$.  Building on this transformation allows us to represent
  the data using a true $d-1$ dimensional distribution, rather than generating a latent parameter
  to induce a $d$ dimensional distribution, as we do with all the gamma based models.

A canonical choice of distribution in this space might be to further transform to $(-\infty, \infty)$
  via marginal probit or logit transformation, and represent the data as multivariate normal.  We try
  that here.   Let $W_i = \text{Probit}(2\theta_i /\pi)$--that is, scale $\theta_i$ to the unit
  interval, then conduct a probit transformation on it to result in $W_i \in (-\infty, \infty)$.
  Then we establish a multivariate normal distribution on $W$. As we are expecting data to descend
  from a mixture of distributions, we place a DP prior on the multivariate normal kernel
  distribution.  The centering distribution of the DP prior is the product of a multivariate normal
  and inverse Wishart distribution; and we place multivarite normal and inverse Wishart priors on
  these parameters.  As before, we place a gamma prior on the DP concentration parameter $\eta$.
  \begin{equation}
    \begin{aligned}
                W_i &\sim \mathcal{N}_{d-1}\left(\mu_i, \Sigma_i\right)\\
    \mu_i, \sigma_i &\sim G_i\\
                G_i &\sim \text{DP}(\eta, G_0(\mu_i,\Sigma_i\mid\mu_0,\Sigma_0))\\
                    &\hspace{1cm}G_0(\mu_i,\Sigma_i\mid\mu_0,\Sigma_0) &=
                      \mathcal{N}_{d-1}(\mu_i\mid\mu_0,\Sigma_0)\text{IW}(\Sigma\mid\nu,\psi)\\
              \mu_0 &\sim \mathcal{N}_{d-1}\left({\bf u},{\bf S}\right)\\
           \Sigma_0 &\sim \text{IW}(\nu_0,\psi_0)\\
               \eta &\sim \text{Ga}(\alpha, \beta)
    \end{aligned}
  \end{equation}
There is an advantage in that for inference on $\mu_i, \mu_0, \Sigma_i$, and $\Sigma_0$ this model
  is completely conjugate.  However, while the transformation employed in Equation~\ref{eqn:transform}
  is one to one, small deviations in different dimensions on $S_{\infty}^{d-1}$ have vastly different
  effects on the resulting transformed variables.  This induced distortion may result in an inferior
  model, when evaluating on $S_{\infty}^{d-1}$.  We will be evaluating this model as representative
  of models on the $d-1$ dimensional coordinate space, and comparing it against other models, after
  projecting back onto $S_{\infty}^{d-1}$.

Another disadvantage of this model is the need to compute $d-1$-dimensional matrix determinants
  and inversions.  If we consider that inversion is a $\mathcal{O}(n^3)$ operation, we face the real
  problem of computation time climbing astronomically as the number of dimensions grows.  We are
  currently evaluating this model on 8 and 46 dimensions, we can see that those operations on 46
  dimensions will take around 190 times longer than on 8 dimensions.  This presents a problem if we
  want to run this model in any reasonable time scale.

% EOF
