\bruno{You said this already}
Assuming that $X_j$ is in the domain of attraction of a max stable distribution for each $j$, then
  the standardization of $X$ to GPD process occurs as follows:

\begin{equation}
Z_{j} = \left(1 + \gamma_j\frac{X_{j} - b_{tj}}{a_{tj}}\right)_{+}^{1/\gamma_j}
\end{equation}
where $b_{tj} = F^{-1}(1-1/t)$, and $a_{tj}$ and $\gamma_{j}$ are evaluated via maximum likelihood.
  The quantity $\left(1 + \gamma_j\frac{X_{j} - b_{tj}}{a_{tj}}\right)_+$ is left truncated at 0.
  Note that $Z_j > 1$ indicates that $X_{j} > b_{tj}$, meaning that the observation was
  \emph{extreme} in that dimension.  Recognize here that $Z_i$ exists on the positive orthant in
  Euclidean space, $\mathcal{R}_+^d$, and that $\max_jZ_j$ follows a simple Pareto distribution.
  Additionally, we can transform ${\bf Z} \rightarrow (R, {\bf V})$, where:
\begin{equation}
  \begin{aligned}
    R &= \lVert {\bf z} \rVert_{\infty} = \max_{j \in (1,\ldots,d)} z_i\\
    {\bf V} &= \left(\frac{z_1}{R},\ldots,\frac{z_d}{R}\right) \in S_{\infty}^{d-1}.\\
  \end{aligned}
\end{equation}
  Thus ${\bf V}$ is the projection of the ${\bf Z}$ vector onto the unit hypersphere defined by
  the $L_{\infty}$ norm, $S_{\infty}^{d-1}$.  As stated before by homogeneity property, a spectral
  measure $\Phi$ on some space in $S_{\infty}^{d-1}$ is independent of $R$, which follows the
  standard Pareto distribution by construction.

We are interested in the angular distribution of ${\bf V}$, the projection of the standardized
  observations ${\bf Z}$ onto $S_{\infty}^{d-1}$.  As we are unaware of any distribution that
  operates natively this space (the positive orthant of the unit hypercube), or can be effectively
  coerced into this space, we construct distributions on other spaces for which there exists a
  one to one mapping from the distribution space and target space.  That means, the space
  $S_{\infty}^{d-1}$ can be projected onto some other $S_p^{d-1}$ using another norm $\mathcal{L}_p$.
  We project onto the unit simplex using the $L_1$ norm, and onto the unit hypersphere using the
  $L_2$ norm.  There exists a one-to-one mapping between these spaces and the unit hypercube.

For model comparison, as the projection induces its own distortion, and owing to the difficulty of
  creating a density directly on $S_{\infty}^{d-1}$, we will conduct model comparison using the
  posterior predictive distributions in $S_{\infty}^{d-1}$, using a scoring rule appropriate
  to that space.

% EOF
